---
title: "Airplane Dataset Analysis"
author: "Group 1 (Harsh Agrawal, Amit Ramjee, Shweta Mishra, Abishek Amberkar)"
date: "12/2/2021"
output:
  pdf_document: default
  word_document: default
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


# Exploratory Analysis - Correlation Heatmap
### Load Training & Test data files, replace blanks with median
[Kaggle Airline Passenger Dataset](https://www.kaggle.com/teejmahal20/airline-passenger-satisfaction)

```{r}
rm(list=ls())
air_training <- read.csv("C:/Users/amitr/Downloads/airplane_train.csv") # Load dataset
air_training$Arrival.Delay.in.Minutes[is.na(air_training$Arrival.Delay.in.Minutes)]<-median(air_training$Arrival.Delay.in.Minutes, na.rm = TRUE)

air_test<- read.csv("C:/Users/amitr/Downloads/airplane_test.csv") # Load dataset
air_test$Arrival.Delay.in.Minutes[is.na(air_test$Arrival.Delay.in.Minutes)]<-median(air_test$Arrival.Delay.in.Minutes, na.rm = TRUE)
```

### Pre-processing, convert categorical columns to numeric for training data

```{r}
# Update categorical columns to numeric for training data
air_training<-air_training[,-1:-2] # Eliminate the first 2 columns for row number and identifier
air_training$Gender<-ifelse(air_training$Gender=="Male", 1, 0)
air_training$Customer.Type<-ifelse(air_training$Customer.Type=="Loyal Customer", 1, 0)
air_training$Type.of.Travel<-ifelse(air_training$Type.of.Travel=="Business travel", 1, 0)
air_training$Business<-ifelse(air_training$Class=="Business", 1, 0)
air_training$EcoPlus<-ifelse(air_training$Class=="Eco Plus", 1, 0)
air_training<-air_training[,-5] # remove Class column since created Business and EcoPlus column from that
air_training$satisfaction<-ifelse(air_training$satisfaction=="satisfied", 1, 0)

```

### Pre-processing, convert categorical columns to numeric for testing data

```{r}
air_test<-air_test[,-1:-2] # Eliminate the first 2 columns for row number and identifier
air_test$Gender<-ifelse(air_test$Gender=="Male", 1, 0)
air_test$Customer.Type<-ifelse(air_test$Customer.Type=="Loyal Customer", 1, 0)
air_test$Type.of.Travel<-ifelse(air_test$Type.of.Travel=="Business travel", 1, 0)
air_test$Business<-ifelse(air_test$Class=="Business", 1, 0)
air_test$EcoPlus<-ifelse(air_test$Class=="Eco Plus", 1, 0)
air_test<-air_test[,-5] # Eliminates class column
air_test$satisfaction<-ifelse(air_test$satisfaction=="satisfied", 1, 0)
```


### Plot a Correlation Matrix with all relevant variables

```{r}
library(ggcorrplot)
dev.off()
cor_matrix<-cor(air_training)
ggcorrplot(cor_matrix, hc.order=TRUE, type = 'lower', lab = TRUE)
```

***
# Exploratory Analysis - How Depature Delayal impact Passenger Satisfaction

```{r}
class(air_training$Departure.Delay.in.Minutes)
ggplot(air_training, aes(x = as.factor(satisfaction), y = Departure.Delay.in.Minutes)) +
  geom_boxplot() +
  coord_cartesian(ylim = quantile(air_training$Departure.Delay.in.Minutes, p = c(0.1, 0.9)))
```

***
# Exploratory Analysis - How do different classes impact Satisfaction (Economy vs. Business)? 
### Prepare data for Satisfied vs. Neutral/Disatisfied in different Classes

```{r}
rm(list=ls())
air_training <- read.csv("C:/Users/amitr/Downloads/airplane_train.csv") # Load dataset
air_training$Arrival.Delay.in.Minutes[is.na(air_training$Arrival.Delay.in.Minutes)]<-median(air_training$Arrival.Delay.in.Minutes, na.rm = TRUE)
air_training<-air_training[,-2] # Eliminate the first 2 columns for row number and identifier
air_training$Gender<-ifelse(air_training$Gender=="Male", 1, 0)
air_training$Customer.Type<-ifelse(air_training$Customer.Type=="Loyal Customer", 1, 0)
air_training$Type.of.Travel<-ifelse(air_training$Type.of.Travel=="Business travel", 1, 0)
air_training$Business<-ifelse(air_training$Class=="Business", 1, 0)
air_training$EcoPlus<-ifelse(air_training$Class=="Eco Plus", 1, 0)
air_training$satisfaction<-ifelse(air_training$satisfaction=="satisfied", 1, 0)
```

## Plot passengers in Different Classes (economy, plus, business)
#### Where 1 = satisfied and 0 = Neutral or disatisfied

```{r}
class(air_training$Class)
ggplot(air_training, aes(x = as.factor(satisfaction), fill = Class)) + 
  geom_bar(position = 'fill')
```

***
## Box Plot of Simliar Passenger variables to compare ratings

```{r}
library(ggpubr)
Food_and_drink<-ggplot(air_training, aes(x = as.factor(satisfaction), y = Food.and.drink)) + 
  geom_boxplot()

seat_comfort<-ggplot(air_training, aes(x = as.factor(satisfaction), y = Seat.comfort)) + 
  geom_boxplot()


Legroom<-ggplot(air_training, aes(x = as.factor(satisfaction), y = Leg.room.service)) + 
  geom_boxplot()

Inflight_enter<-ggplot(air_training, aes(x = as.factor(satisfaction), y = Inflight.entertainment)) + 
  geom_boxplot()

ggarrange(Food_and_drink, seat_comfort, Legroom, Inflight_enter, ncol = 2, nrow = 2)
```

--

***

# TEST 1: Naives Bayes

#### Clear environment - Load training, test data.  Fill blanks with median

```{r}
rm(list=ls())
air_training<- read.csv('C:/Users/amitr/Downloads/airplane_train.csv') # Load Training
air_training$Arrival.Delay.in.Minutes[is.na(air_training$Arrival.Delay.in.Minutes)]<-median(air_training$Arrival.Delay.in.Minutes, na.rm = TRUE)


air_test<- read.csv("C:/Users/amitr/Downloads/airplane_test.csv") # Load Test
air_test$Arrival.Delay.in.Minutes[is.na(air_test$Arrival.Delay.in.Minutes)]<-median(air_test$Arrival.Delay.in.Minutes, na.rm = TRUE)

```


#### Pre-processing, convert categorical columns to numeric for training data
##### Eliminate columns from dataset based on correlation matrix which are Row,
##### ID, Gender, Gate Location, Departure Arrival Time, Arrival Delay in Minutes, Departure delay in minutes

```{r}
air_training<-air_training[,-c(1,2,3,10,12,23,24)] # eliminates columns
air_training$Customer.Type<-ifelse(air_training$Customer.Type=="Loyal Customer", 1, 0)
air_training$Type.of.Travel<-ifelse(air_training$Type.of.Travel=="Business travel", 1, 0)
air_training$Business<-ifelse(air_training$Class=="Business", 1, 0)
air_training$EcoPlus<-ifelse(air_training$Class=="Eco Plus", 1, 0)
air_training<-air_training[,-4] # Eliminates class column since added Business & EcoPlus columns
```

#### Pre-processing, convert categorical columns to numeric for test data

```{r}
air_test<-air_test[,-c(1,2,3,10,12,23,24)] # eliminates columns
air_test$Customer.Type<-ifelse(air_test$Customer.Type=="Loyal Customer", 1, 0)
air_test$Type.of.Travel<-ifelse(air_test$Type.of.Travel=="Business travel", 1, 0)
air_test$Business<-ifelse(air_test$Class=="Business", 1, 0)
air_test$EcoPlus<-ifelse(air_test$Class=="Eco Plus", 1, 0)
air_test<-air_test[,-4] # Eliminates class column

```

#### Factor data

```{r}

air_training$satisfaction <- factor(air_training$satisfaction)
air_test$satisfaction <- factor(air_test$satisfaction)
```

#### Build Naive Bayes model and prediction

```{r}
set.seed(111)
library(e1071)
library(caret)
nB<-naiveBayes(satisfaction~., data = air_training) # this is the model
categorize<-predict(nB, air_test) # this is the prediction
```

#### Return Confusion Matrix & Accuracy
```{r}
table(nB=categorize, Class=air_test$satisfaction)
confusionMatrix(table(air_test$satisfaction, categorize)) # return all stats
```

***

# TEST 2: KNN Method
### This model and all subsequent models use same preprocessing in test 1

#### create vector to store K-values, and variable for error rates

```{r}
k_value<-c(1,3,4,5,6,7,9,12,15,20,30,50,65,80,100) # K value to test
error_Rate<-vector(length = length(k_value))  # store error rate for each model
```

#### run For-loop to return results (25 mins est)

```{r}
library(kknn) # use library to run kknn formula
for(i in 1:length(k_value)){
  knn_model<-kknn(satisfaction~., air_training, air_test,k = k_value[i],kernel = "rectangular")
  predicted<-fitted(knn_model)
  wrong<-sum(predicted!=air_test$satisfaction) # count number of wrong where test data doesn't prediction
  error_Rate[i]<-wrong/length(predicted)
}

print(1-error_Rate) # Print Accuracies
plot(k_value, 1-error_Rate, type = "b", ylab = "Accuracy") # Plot the accuracy
```


***

# TEST 3: CART Method

#### Install Packages for CART

```{r}
library(rpart)
library(rpart.plot)  	
library(rattle)           
library(RColorBrewer)     
library(caret)
```


#### Grow the tree

```{r}
set.seed(111)
CART_class<-rpart(satisfaction~.,data=air_training) # run CART algorithm on training data
CART_class
```

#### Plot the tree

```{r}
dev.off() # reset graphics device
par(mar=c(1,1,1,1))
rpart.plot(CART_class)
```

#### Score the CART

```{r}
CART_predict<-predict(CART_class,air_test, type="class") 
table(Actual=air_test[,"satisfaction"],CART=CART_predict) # reg confusion matrix
str(CART_predict) # check data structure
confusionMatrix(table(actual=air_test$satisfaction,prediction=CART_predict))
```

***

# TEST 4: C5.0 Algorithm

#### Grow the tree

```{r}
library('C50')
C50_class<- C5.0(satisfaction~.,data=air_training)
```

#### Summarize training data - only for viewing purposes

```{r}
summary(C50_class) # this only uses training not test, don't rely on
```

#### Score the C5.0 Model

```{r}
C50_predict<-predict(C50_class, air_test, type="class")
table(actual=air_test[,17],C50=C50_predict)
confusionMatrix(table(actual=air_test[,17],C50=C50_predict))
```


***

# TEST 5: Random Forest Algorithm

#### Set seed, load Random Forest Library

```{r}
set.seed(111)
library(randomForest)
```

#### Create the Random Forest fit model, plot

```{r}
fit <- randomForest(satisfaction~., data=air_training, importance=TRUE, ntree=200)
importance(fit)
varImpPlot(fit)
Prediction <- predict(fit, air_test)
```

#### Score the Random Forest

```{r}
confusionMatrix(table(actual=air_test[,17],Prediction)) #Accuracy results
```


***

# TEST 6: SVM - Support Vector Machine

#### Load SVM package and grow tree

```{r}
library(e1071)
svm.model <- svm( satisfaction~ ., data = air_training  )
svm.pred <- predict(svm.model,  air_test )
```

#### Establish confusion matrix, accuracy

```{r}
table(actual=air_test[,17],svm.pred )
confusionMatrix(table(actual=air_test[,17],svm.pred))
SVM_wrong<- (air_test$satisfaction!=svm.pred)
rate<-sum(SVM_wrong)/length(SVM_wrong)
accuracy = 1-rate
accuracy
```

#### Print

```{r}
print(svm.model)
```

